# Regression and model validation

*During the Data wrangling exercise I created an analysis dataset for the Analysis exercise. During the Analysis exercise I explored the data, performed analysis and interpreted the results.*

*At first, here is the analysis dataset.*
```{r}
learning2014<-read.csv('C:/Users/Henna/OneDrive/Asiakirjat/OPEN SCIENCE/IODS-project/data/learning2014.csv')
learning2014
```

*The dataset has following dimensions..*
```{r}
dim(learning2014)
```
*where at first values is the number of observations and second value is the number of variables(gender, age, attitude, deep, stra, surf and points).*


*Gender: M (Male), F (Female)*
```{r}
gender<-(learning2014$gender)
str(gender)
```
*Structures of the dataset.*

*Variable age (in years) is derived from the date of birth.*
```{r}
age<-learning2014$Age
str(age)
```
*Global attitude toward statistics*
```{r}
attitude<-(learning2014$attitude)
str(attitude)
```
*Deep approach*
```{r}
deep<-(learning2014$deep)
str(deep)
```
*Strategic approach*
```{r}
stra<-(learning2014$stra)
str(stra)
```
*Surface approach*
```{r}
surf<-(learning2014$surf)
str(surf)
```
*Exam points*
```{r}
points<-(learning2014$Points)
str(points)
```

*Graphical overview of the data.*


```{r}
pairs(learning2014[-1], learning2014$gender)

# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(learning2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p


```

*Table plot shows all variables(x,y-axis), measurements(points spread plots) and their correlations graphically(diagonal plots) and with numbers. Upper row in plot shows correlation of gender to other variables as a box plot. The box plot of the data can be generated by calculating five relevant values: minimum, maximum, median, first quartile, and third quartile. Age distribution of collected data is the highest in range of 20-30 years. Distribution on attitude is highest in range of 2.5-3.5(females), 3-3.5(males). Correlations of can be analyzed from the graph. There is strong correlation between age and attitude(0.11), and between deep and stra(0.14). The attitude increases when the age is increasing. This correlation is even stronger with males than females(0.173).*


```{r}
# Use the gglot2 library
library(ggplot2)


# initialize plot with data and aesthetic mapping
p1 <- ggplot(learning2014, aes(x = attitude, y = points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# draw the plot
p2

# add a regression line
p3 <- p2 + geom_smooth(method = "lm")

# add a main title and draw the plot
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4


```

*Student's attitude versus exam points. Variablles gender, point and attitude are  plotted. Regression model fitted, where exam points is the target variable. These tree variables has correlations between each other. The plot shows a summary of the fitted model. Attitude is greater with males and points are higher when the attitude is greater.*
```{r}

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = learning2014) + geom_smooth(method = "lm")

# fit a linear model
my_model <- lm(points ~ attitude, data = learning2014)

# print out a summary of the model
summary(my_model)


```

*Using a summary of your fitted model, explain the relationship between the chosen explanatory variables and the target variable (interpret the model parameters). Explain and interpret the multiple R squared of the model. (0-3 points)*

*Produce the following diagnostic plots: Residuals vs Fitted values, Normal QQ-plot and Residuals vs Leverage. Explain the assumptions of the model and interpret the validity of those assumptions based on the diagnostic plots. (0-3 points)* 


```{r}
date()
```
