---
title: "Clustering and classificatioin "
output: html_document
---

*Boston data was loaded from MASS package.*


```{r}
library(MASS)
# load the data
data("Boston")
# explore the dataset
str(Boston)
dim(Boston)

```
*The Boston data frame has 506 row and 14 columns. Data consist of housing values in suburbs of Boston.*
*Variables are:*

*crim (per capita crime rate by town)*

*zn(proportion of residential land zoned for lots over 25,000 sq.ft)*

*indus(proportion of non-retail business acres per town)*

*chas(Charles River dummy variable (= 1 if tract bounds river; 0 otherwise))*

*nox(nitrogen oxides concentration (parts per 10 million))*

*rm(average number of rooms per dwelling)*

*age(proportion of owner-occupied units built prior to 1940)*

*dis(weighted mean of distances to five Boston employment centres)*

*rad(index of accessibility to radial highways)*

*tax(full-value property-tax rate per \$10,000*

*ptratio(pupil-teacher ratio by town*

*black(1000(Bk - 0.63)^2 where Bk is the proportion of blacks by town*

*lstat(lower status of the population (percent)*

*medv(median value of owner-occupied homes in\$1000s)*


*More details about data can be found from link:* <https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/Boston.html>.


```{r}
library(MASS)
summary(Boston)
# plot matrix of the variables
pairs(Boston)
```

*It is very hard to see the distributions of data and relationships between them, because lots of data and wide range of datapoints. Chas is binary variable, other variables has datapoits in broader distridution. Few varible sets have linear slope, and therefore they have relationship.*

*Few examples of variablees that has relationship between them:*

*medv and lstat*

*nox and dis*

```{r}
# center and standardize variables
boston_scaled <- scale(Boston)
# summaries of the scaled variables
summary(boston_scaled)
# class of the boston_scaled object
class(boston_scaled)
# change the object to data frame
boston_scaled <- as.data.frame(boston_scaled)

```
*Data was scaled so, the mean values is centered to 0 and distribution of variables are in range from +9.9 to - 3.9.*
Create a categorical variable of the crime rate in the Boston dataset (from the scaled crime rate). Use the quantiles as the break points in the categorical variable. Drop the old crime rate variable from the dataset. Divide the dataset to train and test sets, so that 80% of the data belongs to the train set. (0-2 points)

Fit the linear discriminant analysis on the train set. Use the categorical crime rate as the target variable and all the other variables in the dataset as predictor variables. Draw the LDA (bi)plot. (0-3 points)

Save the crime categories from the test set and then remove the categorical crime variable from the test dataset. Then predict the classes with the LDA model on the test data. Cross tabulate the results with the crime categories from the test set. Comment on the results. (0-3 points)

Reload the Boston dataset and standardize the dataset (we did not do this in the Datacamp exercises, but you should scale the variables to get comparable distances). Calculate the distances between the observations. Run k-means algorithm on the dataset. Investigate what is the optimal number of clusters and run the algorithm again. Visualize the clusters (for example with the pairs() or ggpairs() functions, where the clusters are separated with colors) and interpret the results. (0-4 points)




