---
title: "Dimensionality reduction techniques"
output: html_document
---
# Dimensionality reduction techniques

More details about the analyzed data set in link: <http://hdr.undp.org/sites/default/files/hdr2015_technical_notes.pdf>

```{r}
human <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt")
library(ggplot2)
library(GGally)
plot(human)
summary(human)

```




It is very hard to define distributions of the datasets only by looking at the graphiacal overview. There are correlations between variables that show negative or positive slope in graph for example GNI has high relation to all other variables exept to Parli.F.



```{r}
# Access GGally
library(GGally)
ggpairs(human)

```



GGplot shows correlations more clearly. Ado.Birth and Mat.Mor has high positive correlation 0.759 and they have relation. Life.Exp has great positive relation with GNI, Mat.mor. and Ado.Birth.


```{r}
#PCA on the not standardized human data
pca_human <- prcomp(human)
#Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables.
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

```
```{r}
# center and standardize variables
human_scaled <- scale(human)
pca_human_scaled <- prcomp(human_scaled)
#Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables.
biplot(pca_human_scaled, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```



Non-standardized biplot shows counries over the world, standardized biplot centered on the data part from Africa area. Standardized biplot shows dimension that correlate more with Afcica dataset. We can see how data distributes over these dimensions. In standardized plot Labo.FM and Parli F. can be analyzed in PC1 set and rest of the variables in PC2-axis.

The plots shows that countries where is higher education and equality between gender, there is also lower child mortality, maternal mortality and also better life expectancy. 


```{r}
# get all packages available

library(magrittr)
library(ggplot2)
library(tidyr)
library(FactoMineR)

##I had knit problem, so I managed the problem by saving the variable tea into a text file in my data folder. Then I loaded it from there.


tea<-read.csv('C:/Users/Henna/OneDrive/Asiakirjat/OPEN SCIENCE/IODS-project/data/tea.txt')
str(tea)
dim(tea)

# visualize the dataset
gather(tea) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar() + theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 8))

# column names to keep in the dataset
keep_columns <- c("Tea", "How", "how", "sugar", "where", "lunch")

# select the 'keep_columns' to create a new dataset
tea_time <- dplyr::select(tea,one_of(keep_columns))
# multiple correspondence analysis
mca <- MCA(tea_time, graph = FALSE)
# summary of the model
summary(mca)
# visualize MCA
plot(mca, invisible=c("ind"), habillage = "quali")
                                                  
```





The results shows how people drink their tea. For example people who takes green tea, they also  buy tea from chain store. Unpacked teas are bought from tea shop. People who drink black tea they don't use sugar and don't drink it at lunch. People who use tea bag also uses sugar more often.
