# Linear regression

*During the Data wrangling exercise I created an analysis dataset for the Analysis exercise. During the Analysis exercise I explored the data, performed analysis and interpreted the results.*

*At first, here is the analysis of the data set.*
```{r}
learning2014<-read.csv('C:/Users/Henna/OneDrive/Asiakirjat/OPEN SCIENCE/IODS-project/data/learning2014.csv')

```

*The dataset has following dimensions..*
```{r}
dim(learning2014)
```
*where at first values is the number of observations and second value is the number of variables(gender, age, attitude, deep, stra, surf and points).*


*Gender: M (Male), F (Female)*
```{r}
gender<-(learning2014$gender)
str(gender)
```
*Structures of the dataset.*

*Variable age (in years) is derived from the date of birth.*
```{r}
age<-learning2014$Age
str(age)
```
*Global attitude toward statistics*
```{r}
attitude<-(learning2014$attitude)
str(attitude)
```
*Deep approach*
```{r}
deep<-(learning2014$deep)
str(deep)
```
*Strategic approach*
```{r}
stra<-(learning2014$stra)
str(stra)
```
*Surface approach*
```{r}
surf<-(learning2014$surf)
str(surf)
```
*Exam points*
```{r}
points<-(learning2014$Points)
str(points)
```

*Graphical overview of the data.*


```{r}
pairs(learning2014[-1], learning2014$gender)

# access the GGally and ggplot2 libraries
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(learning2014, mapping = aes(col = gender, alpha = 0.3), lower = list(combo = wrap("facethist", bins = 20)))

# draw the plot
p


```

*Table plot shows all variables(x,y-axis), measurements(points spread plots) and their correlations graphically(diagonal plots) and with numbers. Upper row in plot shows correlation of gender to other variables as a box plot. The box plot of the data can be generated by calculating five relevant values: minimum, maximum, median, first quartile, and third quartile. Age distribution of collected data is the highest in range of 20-30 years. Distribution on attitude is highest in range of 2.5-3.5(females), 3-3.5(males). Correlations of can be analyzed from the graph. There is strong correlation between age and attitude(0.11), and between deep and stra(0.14). The attitude increases when the age is increasing. This correlation is even stronger with males than females(0.173).*


```{r}
# Use the gglot2 library
library(ggplot2)


# initialize plot with data and aesthetic mapping
p1 <- ggplot(learning2014, aes(x = attitude, y = points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# draw the plot
p2

# add a regression line
p3 <- p2 + geom_smooth(method = "lm")

# add a main title and draw the plot
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4


```

*Student's attitude versus exam points. Variables gender, point and attitude are  plotted. Regression model is fitted on data points, and where exam points is the target variable. These tree variables has correlations between each other. The plot shows a summary of the fitted model. Attitude is greater with males and the amount of points increases due the attitude is increasing.*
```{r}

# a scatter plot of points versus attitude
library(ggplot2)
qplot(attitude, points, data = learning2014) + geom_smooth(method = "lm")

# fit a linear model
my_model <- lm(points ~ attitude, data = learning2014)

# print out a summary of the model
summary(my_model)


```

*Here is a regression model of relationship between attitude and points. The model is linear regression with positive slope. It means that when the attitude is increasing, also the amount of points is increasing. The magnitude of the slope (regression coefficient) tells the magnitude of the correlation between variables. Since the slope is positive in this model, the variable attitude has positive effect to target variable points.  Explain and interpret the multiple R squared of the model. Valid of the regression model can be estimated by observing how close data points are to the model. More points near by the linear model tells that the model is more accurate. R squared regression model tells how large part of the variability of the target variable can be explained with the other variable. R squared get values between 0 to 1 and can be determined by taking. When R squared is near to one, the model is able to predict the target variables more accurately. In our case the R squared is 0.19, therefore we can come to conclusion that the model can predict the values fairly well*


```{r}
date()
```
