---
title: "Dimensionality reduction techniques"
output: html_document
---


```{r}
human <- read.csv("http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/human2.txt")
library(ggplot2)
library(GGally)
plot(human)
summary(human)

```


It is very hard to define distributions of the datasets only by looking at the graphiacal overview. There are correlations between variables that show negative or positive slope in graph for example GNI has high relation to all other variables exept to Parli.F.
```{r}
# Access GGally
library(GGally)
ggpairs(human)

```
Ado.Birth and Mat.Mor has high positive correlation 0.759 and they have relation. Life.Exp has great positive relation with GNI, Mat.mor. and Ado.Birth.


```{r}
#PCA on the not standardized human data
pca_human <- prcomp(human)
#Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables.
biplot(pca_human, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))

```
```{r}
# center and standardize variables
human_scaled <- scale(human)
pca_human_scaled <- prcomp(human_scaled)
#Draw a biplot displaying the observations by the first two principal components (PC1 coordinate in x-axis, PC2 coordinate in y-axis), along with arrows representing the original variables.
biplot(pca_human_scaled, choices = 1:2, cex = c(0.8, 1), col = c("grey40", "deeppink2"))
```

